---
title: "Q10-Solution"
author: "Sanjeev Gadre"
date: "August 22, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Setting up the environment

```{r environment, message= FALSE, echo=TRUE}
library(ISLR)
library(tidyverse)
library(MASS) #for the LDA and QDA functions
library(class) #for the KNN function
data(Weekly)

```

#10-a
```{r 10-a, echo=TRUE}
names(Weekly)
dim(Weekly)
summary(Weekly)
cor(Weekly[,-9])
Weekly %>% ggplot(aes(Year,Volume, group=Year))+geom_boxplot()
```

The only meaniful correlation is between Year and Volume. The boxplots reveal that <a> the trading volume has increased over years, and <b> so has the weekly variations in the volume in the given year.

#10-b
```{r 10-b, echo=TRUE}
glm_fit = glm(Direction~.-Year-Today, data=Weekly, family=binomial)
summary(glm_fit)
```

Lag2 is the only predictor that appears statistically significant

#10-c
```{r 10-c, echo=TRUE}
glm_probs = predict(glm_fit, data=Weekly, type="response")
glm_pred = rep("Down", 1089)
glm_pred[glm_probs>0.5]="Up"
table(glm_pred, Weekly$Direction)
mean(glm_pred==Weekly$Direction) # % times the prediction is right
```

The accuracy of the logistic predictor is 56.1%. However, the market was "Up" 55.5% (557+48)/1089 of all weeks. So a dumb predictor that predicts "Up" everyday would be right 55.5% times. The logistic predictor built is right 56.1% of the time. The improvement over a dumb predictor is negligible. 

The confusion matrix shows that the logistic predictor overwhelmingly predicts "Up" (90% of the times) and hence does a great job of accurately predicting the days when the market is indeed "Up" (90%), but does a dismal job of predicting the "Down" days (11%).

#10-d
```{r 10-d, echo=TRUE}
train = Weekly$Year%in%1990:2008

Test_Set=Weekly[!train,]
Test_Set_Direction = Weekly$Direction[!train]

glm_fit_1 = glm(Direction~Lag2, data=Weekly, family=binomial, subset=train)
glm_probs_1 = predict(glm_fit_1, Test_Set, type="response")
glm_pred_1 = rep("Down", nrow(Test_Set))
glm_pred_1[glm_probs_1>0.5]="Up"

table(glm_pred_1, Test_Set_Direction)
mean(glm_pred_1==Test_Set_Direction)
```

#10-e
```{r 10-e, echo=TRUE}
lda_fit = lda(Direction~Lag2, data=Weekly, subset=train)
lda_pred = predict(lda_fit, Test_Set)
table(lda_pred$class, Test_Set_Direction)
mean(lda_pred$class==Test_Set_Direction)

```

#10-f
```{r 10-f, echo=TRUE}
qda_fit = qda(Direction~Lag2, data=Weekly, subset=train)
qda_pred = predict(qda_fit, Test_Set)
table(qda_pred$class, Test_Set_Direction)
mean(qda_pred$class==Test_Set_Direction)

```

#10-g
```{r 10-g, echo=TRUE}
train_data = as.matrix(Weekly$Lag2[train])
test_data = as.matrix(Weekly$Lag2[!train])
train_direction = Weekly$Direction[train]

set.seed(1)
knn_pred=knn(train_data,test_data,train_direction,k=1)

table(knn_pred, Test_Set_Direction)
mean(knn_pred==Test_Set_Direction)
```

#Dummy txt. Delete
