---
title: "Q9-Solutions"
author: "Sanjeev Gadre"
date: "September 14, 2018"
output: md_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(boot)
library(MASS)
library(tidyverse)
data("Boston")
```

##9.a & 9.b
```{r 9-a&b}
mu.hat = mean(Boston$medv)
mu.hat*1000

se.hat = sd(Boston$medv)/sqrt(length(Boston$medv))
se.hat*1000

mu.hat.range = c(mu.hat-qnorm(0.975)*se.hat, mu.hat+qnorm(0.975)*se.hat)
mu.hat.range*1000
```

The average of median values of owner-occupied houses across the towns of  Boston is \$22,533 and 95% of these median values lie between \$21,731 and \$23,334.

##9.c
```{r 9-c}
boot.fn = function(data,index){
    mu.hat = mean(data$medv[index])
    return(mu.hat)
}

set.seed(1)
boot(Boston, boot.fn, R=1000)
```

The bootstrap estimate of the standard error in mu.hat is \$412. This compares with \$409 obtained from sample standard deviation in (b) above.

##9.d
```{r 9-d}
c(mu.hat-qnorm(0.975)*0.412, mu.hat+qnorm(0.975)*0.412)*1000
t.test(Boston$medv)
```

The 95% confidence interval for mu.hat based on the bootstrap estimate is \$21,725 - \$23,340. This compares with \$21,730 - \$23,336 obtained using t.test(Boston$medv)

##9.e & 9.f
```{r 9-e&f}
median.hat=median(Boston$medv)
median.hat

boot.median.fn = function(data,index){
  median.hat = median(data$medv[index])
  return(median.hat)
}

boot(Boston, boot.median.fn, R=1000)
```

The median value estimate of the medv in the population is \$21,200 which is slightly lower than the mean value estimate of medv in the population which was \$22,533

The bootstrap estimate for the standard error in the median value of medv in the popuplation is \$391 which compares to the estimate for the standard error in the mean value of medv in the population which was \$412.

##9.g & 9.h
```{r 9-g&h}
mu.hat.10.percentile = quantile(Boston$medv, 0.1)
mu.hat.10.percentile

boot.10.percentile.fn = function(data,index){
  mu.hat.10.percentile = quantile(data$medv[index],0.1)
  return(mu.hat.10.percentile)
}

boot(Boston, boot.10.percentile.fn, R=1000)
```

The bootstrap estimate of the standard error in the tenth percentile of medv is \$513. This compares with the bootstrap estimate of the standard error in mean of medv of \$412. Thus the confidence interval for the tenth percentile of medv is much broader than that for the mean of medv. This stands to reason as the number of observations available in calculating the standard error in the tenth percentile is only a tenth of the number of observations available in calculating the mean and therefore there is greater uncertainity.